Variational Monte Carlo on bosonic systems

### 2018-02-27 project_1/article/article.tex
Fix layout of the full problem

\section{Interacting hard sphere bosons}
    Moving to the full system allowing $\beta$ to vary and setting $a
    \neq 0$ we can write the trial wavefunction as
    \begin{align}
        \Psi_T(\vf{r})
        &=
        \Phi_T(\vf{r})
        J(\vf{r}),
    \end{align}
    where $\ket{\Phi_T}$ is the same Slater permanent as in
    \autoref{eq:initial_trial_wavefunction} and $J(\vf{r})$ is the
    \textit{Jastrow factor} given by
    \begin{align}
        J(\vf{r})
        &=
        \exp\Biggl(
            \sum_{j < l}^N u(r_{jl})
        \Biggr),
    \end{align}
    where $r_{jk} = |\vf{r}_j - \vf{r}_k|$ and
    \begin{align}
        u(r_{jk}) = \ln\bigl[f(a, \vf{r}_j, \vf{r}_k)\bigr].
    \end{align}
            To further shorten the notation we will use $u_{jk} = u(r_{jk})$.
            We wish to find an analytical expression for the local energy.
            Beginning with the gradient we get
            \begin{align}
                \nabla_k\Psi_T(\vf{r})
                &=
                \Bigl[
                    \nabla_k
                    \Phi_T(\vf{r})
                \Bigr]
                J(\vf{r})
                + \Phi_T(\vf{r})
                \nabla_k J(\vf{r}).
            \end{align}
            The gradient of the Slater permament for particle $k$ is given by
            \begin{align}
                \nabla_k
                \Phi_T(\vf{r})
                &=
                \nabla_k\phi_k
                \prod_{i \neq k}^N\phi_i
                = \frac{\nabla_k\phi_k}{\phi_k}
                \Phi_T(\vf{r}).
            \end{align}
            The gradient of the Jastrow factor is given by
            \begin{align}
                \nabla_k J(\vf{r})
                &=
                J(\vf{r})
                \nabla_k\sum_{m < n}^N u_{mn} \\
                &= J(\vf{r})
                \Biggl(
                    \sum_{m = 1}^{k - 1}\nabla_k u_{mk}
                    \sum_{n = k + 1}^N\nabla_k u_{kn}
                \Biggr)
                \\
                &=
                J(\vf{r})
                \sum_{m \neq k}^N\nabla_k u_{km},
            \end{align}
            where the gradient of the interaction term splits the antisymmetric
            sum into two parts. As $r_{ij} = r_{ji}$ we can combine these sums
            into a single sum. This in total yields the gradient
            \begin{align}
                \nabla_k\Psi_T(\vf{r})
                &=
                \Biggl(
                    \frac{\nabla_k\phi_k}{\phi_k}
                    + \sum_{m \neq k}^N
                    \nabla_k u_{km}
                \Biggr)
                \Psi_T(\vf{r}).
                \label{eq:gradient_full_wavefunction}
            \end{align}
            We now find the Laplacian of the trial wavefunction by finding the
            divergence of \autoref{eq:gradient_full_wavefunction}.
            \begin{align}
                \nabla_k^2\Psi_T(\vf{r})
                &=
                \Biggl(
                    \nabla_k\Biggl[
                        \frac{\nabla_k\phi_k}{\phi_k}
                    \Biggr]
                    +
                    \sum_{m \neq k}^N \nabla_k^2 u_{km}
                \Biggr)\Psi_T(\vf{r})
                \\
                &\qquad
                +
                \Biggl(
                    \frac{\nabla_k\phi_k}{\phi_k}
                    + \sum_{m \neq k}^N
                    \nabla_k u_{km}
                \Biggr)^2
                \Psi_T(\vf{r}),
            \end{align}
            where the squared term came from taking the gradient of the trial
            wavefunction.  To further simplify we divide by the trial
            wavefunction. This yields
            \begin{widetext}
                \begin{align}
                    \frac{\nabla_k^2\Psi_T(\vf{r})}{\Psi_T(\vf{r})}
                    &=
                    \Biggl(
                        \nabla_k\Biggl[
                            \frac{\nabla_k\phi_k}{\phi_k}
                        \Biggr]
                        +
                        \sum_{m \neq k}^N \nabla_k^2 u_{km}
                    \Biggr)
                    +
                    \Biggl(
                        \frac{\nabla_k\phi_k}{\phi_k}
                        + \sum_{m \neq k}^N
                        \nabla_k u_{km}
                    \Biggr)^2 \\
                    &=
                    \frac{\nabla_k^2\phi_k}{\phi_k}
                    + 2\frac{\nabla_k\phi_k}{\phi_k}
                    \sum_{m \neq k}\nabla_k u_{km}
                    + \sum_{m\neq k}^N\nabla_k^2 u_{km}
                    + \Biggl(
                        \sum_{m \neq k}^N\nabla_k u_{km}
                    \Biggr)^2.
                \end{align}
            \end{widetext}
            To go from here we have to find the gradient and the Laplacian of
            the single particle functions, $\phi_k$, and the interaction
            functions $u_{km}$. For the single particle functions we use
            Cartesian coordinates when finding the derivatives whereas we for
            the interaction functions will use spherical coordinates and do a
            variable substitution. Beginning with the gradient of the single
            particle functions we get
            \begin{align}
                \nabla_k\phi_k
                &=
                \nabla_k\exp\bigl[
                    -\alpha(x_k^2 + y_k^2 + \beta z_k^2)
                \bigr] \\
                &=
                -2\alpha
                (x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k)
                \phi_k.
            \end{align}
            Note that the subscripts on the unit vectors $\vf{e}_i$ are
            \textit{not} the same as the subscripts used for its components. The
            Laplacian yields
            \begin{align}
                \nabla_k^2\phi_k
                &=
                \Bigl[
                    -2\alpha
                    \big(
                        2 + \beta
                    \bigr)
                    + 4\alpha^2
                    \bigl(
                        x_k^2 + y_k^2 + \beta^2z_k^2
                    \bigr)
                \Bigr]
                \phi_k.
            \end{align}
            In order to derive the interaction functions we have to do a
            variable substitution using $r_{km} = |\vf{r}_k - \vf{r}_m|$. We can
            then rewrite the $\nabla_k$-operator as
            \begin{align}
                \nabla_k
                &=
                \nabla_k
                \dpd[]{r_{km}}{r_{km}}
                =
                \nabla_k r_{km} \dpd[]{}{r_{km}}
                \\
                &=
                \frac{\vf{r}_k - \vf{r}_m}{r_{km}}\dpd[]{}{r_{km}}.
            \end{align}
            Applying this version of the $\nabla_k$-operator to $u_{km}$ yields
            \begin{align}
                \nabla_k u_{km}
                &=
                \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
                \dpd[]{u_{km}}{r_{km}}.
            \end{align}
            For the Laplacian we switch a little back and forth between the two
            ways of representing the $\nabla_k$-operator. We thus get
            \begin{widetext}
                \begin{align}
                    \nabla_k^2 u_{km}
                    &=
                    \frac{\nabla_k \vf{r}_k}{r_{km}}\dpd[]{u_{km}}{r_{km}}
                    + \Biggl[
                        \nabla_k \frac{1}{r_{km}}
                    \Biggr]
                    (\vf{r}_k - \vf{r}_m)\dpd[]{u_{km}}{r_{km}}
                    + \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
                    \nabla_k \dpd[]{u_{km}}{r_{km}} \\
                    &= \frac{d}{r_{km}}\dpd[]{u_{km}}{r_{km}}
                    - \frac{(\vf{r}_k - \vf{r}_m)^2}{r_{km}^3}
                    \dpd[]{u_{km}}{r_{km}}
                    + \frac{(\vf{r}_k - \vf{r}_m)^2}{r_{km}^2}
                    \dpd[2]{u_{km}}{r_{km}} \\
                    &=
                    \frac{d - 1}{r_{km}}\dpd[]{u_{km}}{r_{km}}
                    + \dpd[2]{u_{km}}{r_{km}},
                \end{align}
                where $d$ is again the dimensionality of the problem. In total
                we can state an intermediate version of the Laplacian occuring
                in the local energy as
                \begin{align}
                    \frac{\nabla_k^2\Psi_T(\vf{r})}{\Psi_T(\vf{r})}
                    &=
                    \frac{\nabla_k^2\phi_k}{\phi_k}
                    + 2\frac{\nabla_k\phi_k}{\phi_k}
                    \sum_{m \neq k}^N
                    \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
                    \dpd[]{u_{km}}{r_{km}}
                    + \sum_{m\neq k}^N
                    \Biggl(
                        \frac{d - 1}{r_{km}}\dpd[]{u_{km}}{r_{km}}
                        + \dpd[2]{u_{km}}{r_{km}}
                    \Biggr)
                    \nonumber \\
                    &\qquad
                    +
                    \sum_{m, n \neq k}^N
                    \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
                    \frac{\vf{r}_k - \vf{r}_n}{r_{kn}}
                    \dpd[]{u_{km}}{r_{km}}
                    \dpd[]{u_{kn}}{r_{kn}}.
                \end{align}
            \end{widetext}
            Moving on to the derivatives of the interaction terms, $u_{km}$, to
            get an explicit expression for the Laplacian.
            \begin{align}
                \dpd[]{u_{km}}{r_{km}}
                &=
                \frac{a}{r_{km}(r_{km} - a)},
                \\
                \dpd[2]{u_{km}}{r_{km}}
                &= \frac{a^2 - 2ar_{km}}{r_{km}^2(r_{km} - a)^2}.
            \end{align}

>>>>

\section{Interacting hard sphere bosons}
    Moving to the full system allowing $\beta$ to vary and setting $a
    \neq 0$ we can write the trial wavefunction as
    \begin{align}
        \Psi_T(\vf{r})
        &=
        \Phi_T(\vf{r})
        J(\vf{r}),
    \end{align}
    where $\ket{\Phi_T}$ is the same Slater permanent as in
    \autoref{eq:initial_trial_wavefunction} and $J(\vf{r})$ is the
    \textit{Jastrow factor} given by
    \begin{align}
        J(\vf{r})
        &=
        \exp\Biggl(
            \sum_{j < l}^N u(r_{jl})
        \Biggr),
    \end{align}
    where $r_{jk} = |\vf{r}_j - \vf{r}_k|$ and
    \begin{align}
        u(r_{jk}) = \ln\bigl[f(a, \vf{r}_j, \vf{r}_k)\bigr].
    \end{align}
    To further shorten the notation we will use $u_{jk} = u(r_{jk})$. Computing
    the gradient of the wavefunction we get
    \begin{align}
        \nabla_k\Psi_T(\vf{r})
        &=
        \Bigl[
            \nabla_k
            \Phi_T(\vf{r})
        \Bigr]
        J(\vf{r})
        + \Phi_T(\vf{r})
        \nabla_k J(\vf{r}).
    \end{align}
    The gradient of the Slater permament for particle $k$ is given by
    \begin{align}
        \nabla_k
        \Phi_T(\vf{r})
        &=
        \nabla_k\phi_k
        \prod_{i \neq k}^N\phi_i
        = \frac{\nabla_k\phi_k}{\phi_k}
        \Phi_T(\vf{r}).
    \end{align}
    The gradient of the Jastrow factor is given by
    \begin{align}
        \nabla_k J(\vf{r})
        &=
        J(\vf{r})
        \nabla_k\sum_{m < n}^N u_{mn} \\
        &= J(\vf{r})
        \Biggl(
            \sum_{m = 1}^{k - 1}\nabla_k u_{mk}
            \sum_{n = k + 1}^N\nabla_k u_{kn}
        \Biggr)
        \\
        &=
        J(\vf{r})
        \sum_{m \neq k}^N\nabla_k u_{km},
    \end{align}
    where the gradient of the interaction term splits the antisymmetric
    sum into two parts. As $r_{ij} = r_{ji}$ we can combine these sums
    into a single sum. This in total yields the gradient
    \begin{align}
        \nabla_k\Psi_T(\vf{r})
        &=
        \Biggl(
            \frac{\nabla_k\phi_k}{\phi_k}
            + \sum_{m \neq k}^N
            \nabla_k u_{km}
        \Biggr)
        \Psi_T(\vf{r}).
        \label{eq:gradient_full_wavefunction}
    \end{align}

    The Laplcian of the trial wavefunction is found by finding the divergence of
    \autoref{eq:gradient_full_wavefunction}.
    \begin{align}
        \nabla_k^2\Psi_T(\vf{r})
        &=
        \Biggl(
            \nabla_k\Biggl[
                \frac{\nabla_k\phi_k}{\phi_k}
            \Biggr]
            +
            \sum_{m \neq k}^N \nabla_k^2 u_{km}
        \Biggr)\Psi_T(\vf{r})
        \\
        &\qquad
        +
        \Biggl(
            \frac{\nabla_k\phi_k}{\phi_k}
            + \sum_{m \neq k}^N
            \nabla_k u_{km}
        \Biggr)^2
        \Psi_T(\vf{r}),
    \end{align}
    where the squared term came from taking the gradient of the trial
    wavefunction.  To further simplify we divide by the trial
    wavefunction. This yields
    \begin{align}
        \frac{\nabla_k^2\Psi_T(\vf{r})}{\Psi_T(\vf{r})}
        &=
        \frac{\nabla_k^2\phi_k}{\phi_k}
        + 2\frac{\nabla_k\phi_k}{\phi_k}
        \sum_{m \neq k}\nabla_k u_{km}
        \nonumber \\
        &\qquad
        + \sum_{m\neq k}^N\nabla_k^2 u_{km}
        + \Biggl(
            \sum_{m \neq k}^N\nabla_k u_{km}
        \Biggr)^2.
    \end{align}
    To go from here we have to find the gradient and the Laplacian of
    the single particle functions, $\phi_k$, and the interaction
    functions $u_{km}$. For the single particle functions we use
    Cartesian coordinates when finding the derivatives whereas we for
    the interaction functions will use spherical coordinates and do a
    variable substitution. Beginning with the gradient of the single
    particle functions we get
    \begin{align}
        \nabla_k\phi_k
        &=
        \nabla_k\exp\bigl[
            -\alpha(x_k^2 + y_k^2 + \beta z_k^2)
        \bigr] \\
        &=
        -2\alpha
        (x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k)
        \phi_k.
    \end{align}
    Note that the subscripts on the unit vectors $\vf{e}_i$ are
    \textit{not} the same as the subscripts used for its components. The
    Laplacian yields
    \begin{align}
        \nabla_k^2\phi_k
        &=
        \Bigl[
            -2\alpha
            \big(
                d - 1 + \beta
            \bigr)
            \nonumber \\
            &\qquad
            + 4\alpha^2
            \bigl(
                x_k^2 + y_k^2 + \beta^2z_k^2
            \bigr)
        \Bigr]
        \phi_k,
    \end{align}
    with $d$ as the dimensionality of the problem.
    In order to derive the interaction functions we have to do a
    variable substitution using $r_{km} = |\vf{r}_k - \vf{r}_m|$. We can
    then rewrite the $\nabla_k$-operator as
    \begin{align}
        \nabla_k
        &=
        \nabla_k
        \dpd[]{r_{km}}{r_{km}}
        =
        \nabla_k r_{km} \dpd[]{}{r_{km}}
        \\
        &=
        \frac{\vf{r}_k - \vf{r}_m}{r_{km}}\dpd[]{}{r_{km}}.
    \end{align}
    Applying this version of the $\nabla_k$-operator to $u_{km}$ yields
    \begin{align}
        \nabla_k u_{km}
        &=
        \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
        \dpd[]{u_{km}}{r_{km}}.
    \end{align}

    \begin{widetext}
        For the Laplacian we switch a little back and forth between the
        two ways of representing the $\nabla_k$-operator. We thus get
        \begin{align}
            \nabla_k^2 u_{km}
            &=
            \frac{\nabla_k \vf{r}_k}{r_{km}}\dpd[]{u_{km}}{r_{km}}
            + \Biggl[
                \nabla_k \frac{1}{r_{km}}
            \Biggr]
            (\vf{r}_k - \vf{r}_m)\dpd[]{u_{km}}{r_{km}}
            + \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
            \nabla_k \dpd[]{u_{km}}{r_{km}} \\
            &= \frac{d}{r_{km}}\dpd[]{u_{km}}{r_{km}}
            - \frac{(\vf{r}_k - \vf{r}_m)^2}{r_{km}^3}
            \dpd[]{u_{km}}{r_{km}}
            + \frac{(\vf{r}_k - \vf{r}_m)^2}{r_{km}^2}
            \dpd[2]{u_{km}}{r_{km}} \\
            &=
            \frac{d - 1}{r_{km}}\dpd[]{u_{km}}{r_{km}}
            + \dpd[2]{u_{km}}{r_{km}},
        \end{align}
        where $d$ is again the dimensionality of the problem. In total
        we can state an intermediate version of the Laplacian occuring
        in the local energy as
        \begin{align}
            \frac{\nabla_k^2\Psi_T(\vf{r})}{\Psi_T(\vf{r})}
            &=
            \frac{\nabla_k^2\phi_k}{\phi_k}
            + 2\frac{\nabla_k\phi_k}{\phi_k}
            \sum_{m \neq k}^N
            \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
            \dpd[]{u_{km}}{r_{km}}
            + \sum_{m\neq k}^N
            \Biggl(
                \frac{d - 1}{r_{km}}\dpd[]{u_{km}}{r_{km}}
                + \dpd[2]{u_{km}}{r_{km}}
            \Biggr)
            \nonumber \\
            &\qquad
            +
            \sum_{m, n \neq k}^N
            \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
            \frac{\vf{r}_k - \vf{r}_n}{r_{kn}}
            \dpd[]{u_{km}}{r_{km}}
            \dpd[]{u_{kn}}{r_{kn}}.
        \end{align}
    \end{widetext}
    Moving on to the derivatives of the interaction terms, $u_{km}$, to
    get an explicit expression for the Laplacian.
    \begin{align}
        \dpd[]{u_{km}}{r_{km}}
        &=
        \frac{a}{r_{km}(r_{km} - a)},
        \\
        \dpd[2]{u_{km}}{r_{km}}
        &= \frac{a^2 - 2ar_{km}}{r_{km}^2(r_{km} - a)^2}.
    \end{align}
